Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-05 clustering_infinite_model.tex
typo in intro/motivation and missing alphas in section 4.2
Section 4 done (plus some small changes to Section 1.5)

In this section we will establish the exact expressions for the limit local clustering coefficient and function, $c_\infty$ and $c_{\infty}(k)$, respectively.

We shall first recall some necessary notations for computing the local clustering function in $G_{\Pcal}$ and introduce some new ones.

Recall that for $p\in \R \times \R_+$,
\[
	B(p) = \{p^\prime \in \R \times \R_+ : |x - x^\prime| \le e^{\frac{y + y^\prime}{2}}\}.
\]

Let $T_{\Pcal}(k)$ denote the number of triangles connected to nodes of degree $k$ in $G_{\Pcal}(\alpha, \nu)$, i.e.
\[
	T_{\Pcal}(k) = \sum_{p \in \Pcal} \ind{D_\Pcal(p) = k} T(p) = \sum_{p \in \Pcal} \ind{D_\Pcal(p) = k} \sum_{p_1, p_2 \in 2^\Pcal} T(p,p_1,p_2),
\]
where $2^{\Pcal}$ is the set of distinct pairs in $\Pcal \times \Pcal$ and
\[
	T(p,p_1,p_2) = \ind{p_1 \in B_\Pcal(p)} \ind{p_2 \in B_\Pcal(p)} \ind{p_1 \in B_\Pcal(p_2)},
\]
is the indicator that $\{p, p_1, p_2\}$ form a triangle. Then the local clustering function is given by
\begin{equation}\label{eq:def_local_clustering_P}
	c_{\Pcal}(k) = \begin{cases} 
		\frac{T_{\Pcal}(k)}{\binom{k}{2} N_\Pcal(k)} &\mbox{if } N_\Pcal(k) \ge 1\\
		0 &\mbox{else,}
		\end{cases}
\end{equation}
and the adjusted local clustering by
\begin{equation}\label{eq:def_local_clustering_ast_P}
	c_{\Pcal}^\ast(k) = \begin{cases} 
		\frac{T_{\Pcal}(k)}{\binom{k}{2} \Exp{N_\Pcal(k)}} &\mbox{if } N_\Pcal(k) \ge 1\\
		0 &\mbox{else,}
		\end{cases}.
\end{equation}

To analyze the local clustering we define the conditional expected number of triangles of a vertex $p$ with degree $k$ as
\[
	\Delta_{\Pcal}(p,k) = \CExp{\sum_{(p_1, p_2) \in 2^\Pcal} T_\Pcal(p, p_1, p_2)}{D_\Pcal(p) = k}.
\]
To compute this expression, let $Z_{1}, Z_{2}$ be two independent random variables on $B_{\Pcal}(p)$, with (marginal) probability measures $\eta_{p} = \frac{\mu|_{B_{\Pcal}(p)}}{\mu(B_{\Pcal}(p))}$. Then we define,
\begin{equation}\label{eq:def_delta_p}
	\Delta_{\Pcal}(p) = \Prob{ Z_{1} \in \mathcal{B}_{\Pcal}(Z_{2})} %Z_{1} \in \mathcal{B}_{\Pcal}(p), Z_{2} \in \mathcal{B}_{\Pcal}(p),
\end{equation}
as the probability that two random neighbors of a point $p \in \mathcal{P}$ are adjacent (note that $\pr = \pr_{\mathcal{P},p}$ denotes the joint probability measure of $Z_1$ and $Z_2$).

We shall often abuse notation and write $\Delta_\Pcal(y)$ to denote $\Delta_\Pcal((0,y))$. %Observe that $\Delta_\Pcal(p)$ is the probability that two random neighbors of $p$ are connected.
Next we note that conditioned on the Poisson process $\Pcal$ having $k$ points in $\mathcal{B}_{\Pcal}(p)$, the points are independent random variables $Z_{1}, \dots, Z_{k}$, each with probability measure $\eta_{p}$ and hence
\[
	\Delta_{\Pcal}(p,k) = \binom{k}{2} \Delta_\Pcal(p).
\]

In addition, since $\mu_{\alpha, \nu}(B_\Pcal(p)) = \xi_{\alpha,\nu} e^{\frac{y}{2}}$, with $\xi_{\alpha,\nu} = \frac{4\alpha\nu}{(2\alpha - 1)\pi}$, and
\[
	\Delta_{\Pcal}(p) = \iint_{\Rcal^2} T_\Pcal(p,p_1,p_2) \eta_{p}(x_1,y_1)\eta_{p}(x_2,y_2) \, dx1 \, dx_2 \, dy_1 \, dy_2,
\]
we have that
\begin{equation}\label{eq:relation_T_vs_Delta_P}
	\Exp{T_\Pcal(p)} = \xi_{\alpha,\nu}^2 \frac{e^{y}}{2} \Delta_\Pcal(p).
\end{equation}

\begin{remark}[Notations for the finite graph $G_{\Pcal, n}(\alpha,\nu)$]\label{rmk:notations_clustering_P_n}
For the finite graph $G_{\Pcal,n}(\alpha, \nu)$ we will use notations similar to those for $G_{\Pcal}(\alpha, \nu)$ where we simply introduce a subscript $n$. For example, 
\[
	\Delta_{\Pcal,n}(p) = \pr_{\mathcal{P},p}( Z_{1,n} \in \BallPo{Z_{2,n}}),
\]
%Z_{1,n} \in \BallPo{p}, Z_{2,n} \in \BallPo{p},
where $Z_{1,n}$ and $Z_{2,n}$ are two independent random variables with probability measure \\
$\eta_{p,n}$.
\end{remark}

%and define $\epsilon(p_1,p_2)$ to be the indicator that $p_1$ and $p_2$ are neighbors in $G_{\Pcal}(\alpha,\nu)$, i.e.,
%\[
%	\epsilon(p_1,p_2) = \begin{cases}
%		1 &\mbox{if } p_1 \in B(p_2)\\
%		0 &\mbox{else.}
%	\end{cases}
%\]
%With these definition we then write, 
%\[
%	T_\Pcal(p_1,p_2,p_3) = \epsilon_n(p_1,p_2) \epsilon_n(p_2,p_3) \epsilon_n(p_3,p_1),
%\]
%to denote the indicator that the tripple $(p_1, p_2, p_3)$ forms a triangle in $G_\Pcal(\alpha,\nu)$ and the triangle counting function as
%\begin{equation}
%	T_{\Pcal}(p) = \sum_{p_1, p_2 \in \Pcal} 
%		\hspace{-5pt} T_\Pcal(p,p_1,p_2).
%\end{equation}
%
%The local clustering function then is
%\begin{equation}\label{eq:def_local_clustering_P}
%	c_{\Pcal}(k) = \begin{cases} 
%		\frac{1}{N_\Pcal(k)}\sum_{p \in \Pcal} \ind{D_\Pcal(p) = k} \frac{T_{\Pcal}(k)}{\binom{k}{2}} &\mbox{if } N_k \ge 1\\
%		0 &\mbox{else,}
%		\end{cases}
%\end{equation}
%and similarly for $c_{\Pcal}^\ast(k)$.
%
%To analyze the local clustering we define the conditional expectation of triangles of a vertex $p$ with degree $k$ as
%\[
%	\Delta_{\Pcal}(p,k) = \CExp{\sum_{(p_1, p_2) \in \Pcal} T_\Pcal(p, p_1, p_2)}{p, D_\Pcal(p) = k}.
%\]
%To compute this expression, let $Z_{1}, Z_{2}$ be two independent random variables on $B_{\Pcal}(p)$, with probability measure $\eta_{p} = \frac{\mu|_{B_{\Pcal}(p)}}{\mu(B_{\Pcal}(p))}$. Then we define,
%\begin{equation}\label{eq:def_delta_p}
%	\Delta_{\Pcal}(p) = \CProb{Z_{1} \in B_{\Pcal}(p), Z_{2} \in B_{\Pcal}(p), Z_{1} \in B_{\Pcal}(Z_{2})}{p}
%\end{equation}
%We shall often abuse notation and write $\Delta_\Pcal(y)$ to denote $\Delta_\Pcal((0,y))$. Observe that $\Delta_\Pcal(p)$ is the probability that two random neighbors of $p$ are connected. Next we note that conditioned on the Poisson process $\Pcal$ having $k$ points in $B_{\Pcal}(p)$, the points are independent random variables $Z_{1}, \dots, Z_{k}$ with probability $\eta_{p}$ and hence
%\[
%	\Delta_{\Pcal}(p,k) = \binom{k}{2} \Delta_\Pcal(p).
%\]
%
%\begin{remark}[Notations for the finite graph $G_{\Pcal, n}(\alpha,\nu)$]\label{rmk:notations_clustering_P_n}
%For the finite graph $G_{\Pcal,n}(\alpha, \nu)$ we will use notations similar to those for $G_{\Pcal}(\alpha, \nu)$ where we simply introduce a subscript $n$. For example, 
%\[
%	\Delta_{\Pcal,n}(p) = \CProb{Z_{1,n} \in \BallPo{p}, Z_{2,n} \in \BallPo{p}, Z_{1,n} \in \BallPo{Z_{2,n}}}{p},
%\]
%where $Z_{1,n}$ and $Z_{2,n}$ are two independent random variables with probability measure \\
%$\eta_{p,n} = \mu_n|_{\BallPo{p}}/\mu_n(\BallPo{p})$.
%\end{remark}

We will start with some preliminary results on the degrees in $G_\Pcal(\alpha, \nu)$. The proof of Theorem \ref{thm:asymptotics_average_clustering_P} can be found in Section \ref{ssec:asymptotics_local_clustering_P}. The key ingredient for the proof is a result concerning the asymptotic behavior of $\Delta_\Pcal(y)$ (Proposition \ref{prop:asymptotics_Delta_y_P}) whose proof can be found in Section \ref{ssec:asymptotics_Delta_y_P}.

@@

We only give the proof for the case $\frac{1}{2} < \alpha < \frac{3}{4}$. The proofs for the other two cases are similar. Let $\varphi_1(y)$ be defined as in Proposition \ref{prop:asymptotics_Delta_y_P} and define
\[
	\epsilon_1(k) = \int_0^{\infty} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y
\]
Then we have that
\[
	\int_0^{\infty} \Delta_{\mathcal{P}}(y) \rho(y,k) e^{-\alpha y} \dd y
	= C_\alpha \int_0^{\infty} e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y + \epsilon_1(k).
\]

Using \eqref{eq:general_integral_rho_y_k} with $\beta = 2\alpha - 1$ yields,
\begin{equation}\label{eq:asymptotics_clustering_integral}
	\int_0^{\infty} e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y 
	\sim \xi_{\alpha,\nu}^{6\alpha - 2} k^{-(6\alpha - 1)},
\end{equation}
as $k \to \infty$. Therefore, by applying the asymptotic scaling for the denominator~\eqref{eq:degree_distribution_P_asymptotics} we get
\begin{align*}
	c_\infty(k) &= \frac{\int_0^{\infty} \Delta_{\mathcal{P}}(y) \rho(y,k) e^{-\alpha y} \dd y}
		{\int_0^\infty \rho_{y}(k) e^{-\alpha y} \dd y} \\
	&\sim C_\alpha \xi_{\alpha,\nu}^{4\alpha - 2} \frac{k^{-(6\alpha - 1)}}{k^{-(2\alpha - 1)}}
	+ \frac{\epsilon_1(k)}{\xi_{\alpha,\nu}^{2\alpha} k^{-(2\alpha - 1)}}\\
	&= C_\alpha \xi_{\alpha,\nu}^{4\alpha - 2} k^{-4\alpha + 2} 
    + \xi_{\alpha,\nu}^{-2\alpha} k^{2\alpha - 1}\epsilon_1(k).
\end{align*}
Hence, to prove the result it is enough to show that
\[
	\lim_{k \to \infty} k^{6\alpha - 3}\epsilon_1(k) = 0.
\]
For this we fix $0 < \varphi < 1$, define $a(k) = 2\log((1-\varphi)k)$ and split the integral into two pieces
\begin{align*}
	\epsilon_1(k) &= \int_0^{\infty} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y\\
	&= \int_0^{a(k)} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y\\
	&\hspace{10pt}+ \int_{a(k)}^{\infty} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} 
    	\rho(y,k) e^{-\alpha y} \dd y.
\end{align*} 
For the first integral we use that $\rho(y,k)$ is increasing on $[0,a(k)]$ and
\[
	k^{6\alpha - 3}\rho(a(k),k) = O\left(\frac{k^{6\alpha - 3 + (1-\varphi)k}}{k!} e^{-(1-\varphi)k}\right) = o(1),
\]
to obtain
\begin{align*}
	&\hspace{-30pt}\lim_{k \to \infty} k^{6\alpha - 3} \int_0^{a_n} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} 
    	\rho(y,k) e^{-\alpha y} \dd y\\
	&\le \lim_{n \to \infty} \left(\max_{0 \le y \le a(k)} \varphi_1(y)\right) k^{6\alpha - 1} 
    	\rho(a(k), k) \int_0^\infty e^{-\alpha y} \dd y = 0.
\end{align*}
For the second integral we first note that due to \eqref{eq:asymptotics_clustering_integral}
\[
	\int_0^\infty e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y = \bigT{k^{-(6\alpha - 1)}}.
\] 
Therefore, using that $\varphi_1(y)$ is eventually decreasing and $\lim_{k \to \infty} \varphi_1(a(k)) = 0$, we obtain
\begin{align*}
	&\hspace{-30pt}\lim_{k \to \infty} k^{6\alpha - 3} \int_{a(k)}^{\infty} \varphi_1(y) 
    	e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y\\
	&\le \lim_{k \to \infty} \varphi_1(a(k)) k^{6\alpha - 3} \int_0^\infty 
    	e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) e^{-\alpha y} \dd y = 0,
\end{align*}
which finishes the proof.
\end{proof}

@@

For notational convenience we will switch to using $y_0$ instead of $y$. Note that the random variables $Z_i = (x_i,y_i)$, for $i = 1,2$, correspond to first sampling $y_i$ according to the density
\[
	g(y) := \left(\alpha -\frac{1}{2}\right) e^{-(\alpha-\frac{1}{2})y_1}
\] 
and then sampling $x_i$ uniformly in $[-e^{\frac{1}{2}(y_0+y_1)},e^{\frac{1}{2}(y_0+y_1)}]$. With this in mind we define $P(y_0,y_1,y_2)$ to be the probability that $(0,y_0), (x_1,y_1), (x_2,y_2)$ form a triangle where $x_1$ and $x_2$ are uniform random variables in, respectively, $[-e^{\frac{1}{2}(y_0+y_1)},e^{\frac{1}{2}(y_0+y_1)}]$ and  $[-e^{\frac{1}{2}(y_0+y_2)},e^{\frac{1}{2}(y_0+y_2)}]$.

@@

\paragraph{When $\bm{\alpha \notin \{3/4,1\}}$}
Note that when writing $\Delta_{\Pcal}(y_0)$ as in integral as in equation \eqref{eq:def_delta_p}, by symmetry in the integration variables $y_1$ and $y_2$, we can assume that $y_1<y_2$ in which case either $y_0$ or $y_1$ is the smallest height. This gives half the value of $\Delta_{\Pcal}(y_0)$ and hence
\[ 
	\Delta_\Pcal(y_0) = 2(I_1(y_0) +I_2(y_0)), 
\] 
where $I_1$ and $I_2$ are given by:
\begin{align*}
	I_1(y_0) &:= \int_{0<y_0<y_1<y_2} P(y_0,y_1,y_2) \cdot (\alpha-1/2)^2 e^{-(\alpha-1/2)(y_1+y_2)}  {\dd} y_2 {\dd} y_1 \\ 
	I_2(y_0) &:= \int_{0<y_1<y_0,y_2} P(y_0,y_1,y_2) \cdot (\alpha-1/2)^2 e^{-(\alpha-1/2)(y_1+y_2)} {\dd} y_2 {\dd} y_1 \\ 
\end{align*}

@@

We want to compute the limit $\lim_{\alpha \rightarrow 1} \Delta_{\Pcal_{\alpha,\nu}}(y_0(z_0))$. For this, after factoring out $\frac{1}{\alpha-1}$ we compute the first two terms of the Taylor expansion of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$ in $\alpha$ based at $\alpha = 1$. The constant term of this Taylor expansion is verified to vanish whereas the linear term gives the expression searched (after cancellation with $\frac{1}{\alpha-1}$). Note that two terms of $P(y_0(z_0))$ contain $(\alpha-1)^2$ in their denominator, so for those two another $\frac{1}{\alpha-1}$ is factored out and the first three terms of the Taylor expansion are computed. Their constant terms cancel and after multiplication with $\frac{1}{\alpha-1}$ their linear and quadratic terms become part of the constant and linear term of the overall expansion.

>>>>

In this section we will establish the exact expressions for the limit local clustering coefficient and function, $c_\Pcal$ and $c_{\Pcal}(k)$, respectively. This is done in several steps. Recall that for $p\in \R \times \R_+ := \Rcal$,
\[
	\BallPo{p} = \{p^\prime \in \R \times \R_+ : |x - x^\prime| \le e^{\frac{y + y^\prime}{2}}\},
\]
denotes the neighborhood ball of $p$. Furthermore we recall the definition of $\Delta_\Pcal$ 
\[
	\Delta_{\Pcal}(y) = \iint_{\Rcal^2} T_\Pcal(p,p_1,p_2) \eta_{y}(x_1,y_1)\eta_{y}(x_2,y_2) \, dx1 \, dx_2 \, dy_1 \, dy_2,
\]
where $\eta_{y} = \frac{\mu_{\alpha,\nu}|_{\BallPo{(0,y)}}}{\Mu{\BallPo{(0,y)}}}$ is the marginal probability density associated with a randomly sampled neighbor of $(0,y)$.

We already noted that $\Delta_\Pcal(y)$ denotes the probability that two randomly sampled neighbors of $(0,y)$ are connected. More precisely, let $Z_{1}, Z_{2}$ be two independent random variables on $\BallPo{(0,y)}$, with (marginal) probability measures $\eta_{y}$. Then
\begin{equation}\label{eq:def_delta_p_alt}
	\Delta_{\Pcal}(p) = \pr_{\Pcal,p}\left( Z_{1} \in \mathcal{B}_{\Pcal}(Z_{2})\right)
\end{equation}
where $\pr_{\Pcal,p}$ denotes the joint probability measure of $Z_1$ and $Z_2$. We shall often abuse notation and write $\Delta_\Pcal(y)$ to denote $\Delta_\Pcal((0,y))$.

%Next we note that conditioned on the Poisson process $\Pcal$ having $k$ points in $\mathcal{B}_{\Pcal}(p)$, the points are independent random variables $Z_{1}, \dots, Z_{k}$, each with probability measure $\eta_{p}$ and hence
%\[
%	\Delta_{\Pcal}(p,k) = \binom{k}{2} \Delta_\Pcal(p).
%\]
%
%In addition, since $\mu_{\alpha, \nu}(B_\Pcal(p)) = \xi_{\alpha,\nu} e^{\frac{y}{2}}$, with $\xi_{\alpha,\nu} = \frac{4\alpha\nu}{(2\alpha - 1)\pi}$, and
%\[
%	\Delta_{\Pcal}(p) = \iint_{\Rcal^2} T_\Pcal(p,p_1,p_2) \eta_{p}(x_1,y_1)\eta_{p}(x_2,y_2) \, dx1 \, dx_2 \, dy_1 \, dy_2,
%\]
%we have that
%\begin{equation}\label{eq:relation_T_vs_Delta_P}
%	\Exp{T_\Pcal(p)} = \xi_{\alpha,\nu}^2 \frac{e^{y}}{2} \Delta_\Pcal(p).
%\end{equation}

We will start with some preliminary results on the degrees in $G_\Pcal(\alpha, \nu)$. The proof of Theorem \ref{thm:asymptotics_average_clustering_P} can be found in Section \ref{ssec:asymptotics_local_clustering_P}. The key ingredient for the proof is a result concerning the asymptotic behavior of $\Delta_\Pcal(y)$ (Proposition \ref{prop:asymptotics_Delta_y_P}) whose proof can be found in Section \ref{ssec:asymptotics_Delta_y_P}. The final computations for the exact expressions of $c_\infty$ in Theorem~\ref{thm:clustering_coefficient_hyperbolic} and $c_\infty(k)$ in Theorem~\ref{thm:local_clustering_hyperbolic} can be found in Section~\ref{ssec:exact_expressions_clustering_P}.

@@

We only give the proof for the case $\frac{1}{2} < \alpha < \frac{3}{4}$. The proofs for the other two cases are similar. Let $\varphi_1(y)$ be defined as in Proposition \ref{prop:asymptotics_Delta_y_P} and define
\[
	\epsilon_1(k) = \int_0^{\infty} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y
\]
Then we have that
\[
	\int_0^{\infty} \Delta_{\mathcal{P}}(y) \rho(y,k) \alpha e^{-\alpha y} \dd y
	= C_\alpha \int_0^{\infty} e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y + \epsilon_1(k).
\]

Using \eqref{eq:general_integral_rho_y_k} with $\beta = 2\alpha - 1$ yields,
\begin{equation}\label{eq:asymptotics_clustering_integral}
	\int_0^{\infty} e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y 
	\sim 2\alpha \xi_{\alpha,\nu}^{6\alpha - 2} k^{-(6\alpha - 1)},
\end{equation}
as $k \to \infty$. Therefore, by applying the asymptotic scaling for the denominator~\eqref{eq:degree_distribution_P_asymptotics} we get
\begin{align*}
	c_\infty(k) &= \frac{\int_0^{\infty} \Delta_{\mathcal{P}}(y) \rho(y,k) \alpha  e^{-\alpha y} \dd y}
		{\int_0^\infty \rho_{y}(k) \alpha e^{-\alpha y} \dd y} \\
	&\sim C_\alpha 2\alpha \xi_{\alpha,\nu}^{4\alpha - 2} \frac{k^{-(6\alpha - 1)}}{k^{-(2\alpha - 1)}}
	+ \frac{\epsilon_1(k)}{2\alpha \xi_{\alpha,\nu}^{2\alpha} k^{-(2\alpha - 1)}}\\
	&= C_\alpha 2\alpha \xi_{\alpha,\nu}^{4\alpha - 2} k^{-4\alpha + 2} 
    + \frac{1}{2\alpha}\xi_{\alpha,\nu}^{-2\alpha} k^{2\alpha - 1}\epsilon_1(k).
\end{align*}
Hence, to prove the result it is enough to show that
\[
	\lim_{k \to \infty} k^{6\alpha - 3}\epsilon_1(k) = 0.
\]
For this we fix $0 < \varphi < 1$, define $a(k) = 2\log((1-\varphi)k)$ and split the integral into two pieces
\begin{align*}
	\epsilon_1(k) &= \int_0^{\infty} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y\\
	&= \int_0^{a(k)} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y\\
	&\hspace{10pt}+ \int_{a(k)}^{\infty} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} 
    	\rho(y,k) \alpha e^{-\alpha y} \dd y.
\end{align*} 
For the first integral we use that $\rho(y,k)$ is increasing on $[0,a(k)]$ and
\[
	k^{6\alpha - 3}\rho(a(k),k) = O\left(\frac{k^{6\alpha - 3 + (1-\varphi)k}}{k!} e^{-(1-\varphi)k}\right) = o(1),
\]
to obtain
\begin{align*}
	&\hspace{-30pt}\lim_{k \to \infty} k^{6\alpha - 3} \int_0^{a_n} \varphi_1(y) e^{-(4\alpha - 2)\frac{y}{2}} 
    	\rho(y,k) \alpha e^{-\alpha y} \dd y\\
	&\le \lim_{n \to \infty} \left(\max_{0 \le y \le a(k)} \varphi_1(y)\right) k^{6\alpha - 1} 
    	\rho(a(k), k) \int_0^\infty \alpha e^{-\alpha y} \dd y = 0.
\end{align*}
For the second integral we first note that due to \eqref{eq:asymptotics_clustering_integral}
\[
	\int_0^\infty e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y = \bigT{k^{-(6\alpha - 1)}}.
\] 
Therefore, using that $\varphi_1(y)$ is eventually decreasing and $\lim_{k \to \infty} \varphi_1(a(k)) = 0$, we obtain
\begin{align*}
	&\hspace{-30pt}\lim_{k \to \infty} k^{6\alpha - 3} \int_{a(k)}^{\infty} \varphi_1(y) 
    	e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y\\
	&\le \lim_{k \to \infty} \varphi_1(a(k)) k^{6\alpha - 3} \int_0^\infty 
    	e^{-(4\alpha - 2)\frac{y}{2}} \rho(y,k) \alpha e^{-\alpha y} \dd y = 0,
\end{align*}
which finishes the proof.
\end{proof}

@@

For notational convenience we will switch to using $y_0$ instead of $y$. Note that the random variables $Z_i = (x_i,y_i)$, for $i = 1,2$, correspond to first sampling $y_i$ according to the density
\[
	g(y_i) := \left(\alpha -\frac{1}{2}\right) e^{-(\alpha-\frac{1}{2})y_i}
\] 
and then sampling $x_i$ uniformly in $[-e^{\frac{1}{2}(y+y_i)},e^{\frac{1}{2}(y+y_i)}]$. With this in mind we define $P(y_0,y_1,y_2)$ to be the probability that $(0,y_0), (x_1,y_1), (x_2,y_2)$ form a triangle where $x_1$ and $x_2$ are uniform random variables in, respectively, $[-e^{\frac{1}{2}(y_0+y_1)},e^{\frac{1}{2}(y_0+y_1)}]$ and  $[-e^{\frac{1}{2}(y_0+y_2)},e^{\frac{1}{2}(y_0+y_2)}]$.

@@

\paragraph{When $\bm{\alpha \notin \{3/4,1\}}$}
Note that when writing $\Delta_{\Pcal}(y_0)$ as in integral as in equation \eqref{eq:def_delta_p}, by symmetry in the integration variables $y_1$ and $y_2$, we can assume that $y_1<y_2$ in which case either $y_0$ or $y_1$ is the smallest height. This gives half the value of $\Delta_{\Pcal}(y_0)$ and hence
\[ 
	\Delta_\Pcal(y_0) = 2(I_1(y_0) +I_2(y_0)), 
\] 
where $I_1$ and $I_2$ are given by:
\begin{align*}
	I_1(y_0) &:= \int_{0<y_0<y_1<y_2} P(y_0,y_1,y_2) \cdot (\alpha-1/2)^2 e^{-(\alpha-1/2)(y_1+y_2)}  \dd y_2 \dd y_1 \\ 
	I_2(y_0) &:= \int_{0<y_1<y_0,y_2} P(y_0,y_1,y_2) \cdot (\alpha-1/2)^2 e^{-(\alpha-1/2)(y_1+y_2)} \dd y_2 \dd y_1 \\ 
\end{align*}

@@

We want to compute the limit $\lim_{\alpha \rightarrow 1} \Delta_{\Pcal_{\alpha,\nu}}(y_0(z_0))$. For this, after factoring out $\frac{1}{\alpha-1}$ in the expression for $\Delta_{\Pcal}(y)$ we obtained for $\alpha \ne 1$, we compute the first two terms of the Taylor expansion of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$ in $\alpha$ based at $\alpha = 1$. The constant term of this Taylor expansion is verified to vanish whereas the linear term gives the expression searched (after cancellation with $\frac{1}{\alpha-1}$). Note that two terms of $P(y_0(z_0))$ contain $(\alpha-1)^2$ in their denominator, so for those two another $\frac{1}{\alpha-1}$ is factored out and the first three terms of the Taylor expansion are computed. Their constant terms cancel and after multiplication with $\frac{1}{\alpha-1}$ their linear and quadratic terms become part of the constant and linear term of the overall expansion.

The first summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$ is,
\begin{align*}
	S_1 = -\frac{1}{8 \alpha} = -\frac{1}{8}+\frac{1}{8}(\alpha-1) +O( (\alpha-1)^2)
\end{align*}
The second summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$,
\begin{align*}
	S_2 = (\alpha-1/2)z_0 = \frac{1}{2}z_0 +z_0 (\alpha-1)+O( (\alpha-1)^2)
\end{align*}
The third summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$, needs a 3-term Taylor expansion after another multiplication with $(\alpha-1)$,
\begin{align*}
	S_3 = - \frac{(\alpha - 1/2)^2 z_0^2}{4} = -\frac{1}{16}z_0^2 -\frac{1}{4}z_0^2(\alpha-1)-\frac{1}{4}z_0^2(\alpha-1)^2+O( (\alpha-1)^3)
\end{align*}
The fourth summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$, needs a  3-term Taylor expansion after another multiplication with $(\alpha-1)$,
\begin{align*}
	S_4 = z_0^{-2 + 4 \alpha} \frac{2^{-4 \alpha-1} (3 \alpha - 1)}{\alpha} &= \frac{1}{16}z_0^2 +\frac{z_0^2}{4}\left(\frac{1}{8}+\ln\frac{z_0}{2}\right)(\alpha-1)\\
&\qquad+\frac{z_0^2}{8}\left(4\left(\ln\frac{z_0}{2}\right)^2+\ln\frac{z_0}{2} - \frac{1}{4}\right)(\alpha-1)^2+O( (\alpha-1)^3)
\end{align*}
as
\begin{align*}
\frac{d}{d\alpha}\left( \left(\frac{z_0}{2}\right)^{-2+4\alpha} 2^{-3}\left(3-\frac{1}{\alpha}\right) \right) = \left(\frac{z_0}{2}\right)^{-2+4\alpha}2^{-3}\left(4\ln\left(\frac{z_0}{2}\right)\left(3-\frac{1}{\alpha}\right)+\frac{1}{\alpha^2}\right)
\end{align*}
and
\begin{align*}
\frac{d^2}{d\alpha^2}\left( \left(\frac{z_0}{2}\right)^{-2+4\alpha} 2^{-3}\left(3-\frac{1}{\alpha}\right) \right) = \left( \frac{z_0}{2} \right)^{-2+4\alpha}2^{-3}\left(16\left(\ln\frac{z_0}{2}\right)^2\left(3-\frac{1}{\alpha}\right)+\frac{8\ln(\frac{z_0}{2})}{\alpha^2}-\frac{2}{\alpha^3} \right)
\end{align*}
For the fifth summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$ we get,
\begin{align*}
	S_5 &= z_0^{-2 + 4 \alpha} \frac{(\alpha - 1/2 ) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)}{2\alpha} \\
	&= \frac{z_0^2}{4} B^-(1/2;3,0) + z_0^2\left(\left(\ln(z_0)+\frac{1}{4}\right) B^-(1/2;3,0) \right. \\
 	&\hspace{20pt}\left.+1/2\int_0^{\frac{1}{2}} \ln(t)t^2(1-t)^{-1}+\ln(1-t)t^2(1-t)^{-1}dt \right)  (\alpha-1)+O( (\alpha-1)^2 )
\end{align*}
as
\begin{align*}
	&\hspace{-20pt}\frac{d}{d \alpha}\left( z_0^{-2 + 4 \alpha} \frac{(\alpha - 1/2 )}{2\alpha} 
		B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\right) \\
	&= \ln(z_0) 4 z_0^{-2+4\alpha} \left(\frac{1}{2}-\frac{1}{4\alpha}\right)B^-(1/2;1+2\alpha,-2+2\alpha)\\
	&+z_0^{-2+4\alpha}\frac{1}{4\alpha^2}B^-(1/2;1+2\alpha,-2+2\alpha)+z_0^{-2+4\alpha}\left(\frac{1}{2}-\frac{1}{4\alpha}\right)\int_0^{\frac{1}{2}} \frac{d}{d\alpha}(t^{2\alpha}(1-t)^{2\alpha-3})dt
\end{align*}
where
\begin{align*}
\int_0^{\frac{1}{2}} \frac{d}{d\alpha}(t^{2\alpha}(1-t)^{2\alpha-3})dt = \int_0^{\frac{1}{2}} \ln(t)2t^{2\alpha}(1-t)^{2\alpha-3}+\ln(1-t)2t^{2\alpha}(1-t)^{2\alpha-3}dt
\end{align*}
The sixth summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$,
\begin{align*}
	S_6 = \frac{(1 - z_0)^{2 \alpha}}{8 \alpha} = \frac{(1-z_0)^2}{8}+\frac{(1-z_0)^2}{4} (\ln(1-z_0)-1/2 )(\alpha-1)+O( (\alpha-1)^2 )
\end{align*}
and finally the seventh and last summand of $\Delta_\Pcal(y_0(z_0))(\alpha-1)$ gives,
\begin{align*}
	S_7 = - \frac{z_0^{4 \alpha - 2} B^-(1 - z_0; 2 \alpha, 3 - 4 \alpha)}{4} = -\frac{z_0^2}{4}B^-(1-z_0;2,-1) - z_0^2 \left(\ln(z_0)B^-(1-z_0;2,-1) \right. \\
\left.+\int_0^{1-z_0} 1/2\ln(t)t(1-t)^{-2}-t\ln(1-t)(1-t)^{-2}dt \right) (\alpha-1)+O((\alpha-1)^2)
\end{align*}

When adding these terms we see that the constant coefficients cancel:
